---
title: "Visualize Data"
author: "Sam Minard"
format: gfm
---


```{r}
#| label: setup
#| include: false
library(tidyverse)
library(ggplot2)
```



## Your Turn 0

Add a setup chunk that loads the tidyverse packages.

```{r}
mpg
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```


## Your Turn 2

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = (fct_reorder(class, hwy)),
                             y = hwy)) 
  
  
```


## Your Turn 3

Make a histogram of the `hwy` variable from `mpg`. Hint: do not supply a y variable.

```{r}
ggplot(mpg) +
  geom_histogram(aes(x = hwy))
```

## Your Turn 4

Use the help page for `geom_histogram` to make the bins 2 units wide.

```{r}
ggplot(mpg) +
  geom_histogram(aes(x = hwy), binwidth = 2)
```


## Your Turn 5

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  labs(x = "Engine Size Displacement",
       y = "Highway Mileage",
       color = "Class of Vehicle",
       title = "The larger the engine, the lower the highway mileage") +
  theme_light()
```

## Help Me

What do `facet_grid()` and `facet_wrap()` do? 
(run the code, interpret, convince your group)

```{r}
# Makes a plot that the commands below will modify
q <- ggplot(mpg) + geom_point(aes(x = displ, y = hwy))

#makes 1 row, 4 cols by cyl
q + facet_grid(. ~ cyl)

#makes 3 rows, 1 col by drv
q + facet_grid(drv ~ .)

#makes 3 rows by drv, 4 cols by cyl
q + facet_grid(drv ~ cyl)

#makes a graph for each value in class
q + facet_wrap(~ class)
```


## Your Turn 6

Make a bar chart `class` colored by `class`. Use the help page for `geom_bar` to choose a "color" aesthetic for class.

```{r}
ggplot(mpg) +
  geom_bar(aes(x = class, 
               fill = class)) +
  labs(x = "Class of Vehicle",
       y = "Number of Vehicles in Sample") +
  guides(fill = "none")
theme_bw()

ggplot(mpg) +
  geom_bar(aes(x = class, 
               fill = drv),
           position = "dodge") +
  labs(x = "Class of Vehicle",
       y = "Number of Vehicles in Sample",
       fill = "Type of Drive") +
  theme_bw()
```

## Quiz

What will this code do?

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(color = "black", 
              se = FALSE, 
              method = lm) +
  scale_color_viridis_d()
ggsave("example.jpg", width = 6, height = 4)
```


***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```